---
title: "Chapter 4 Answers"
geometry: margin=1.0in
---
\pagenumbering{gobble}

# Exercises, Chapter 4

__4.1__ Use the ```rep``` function to create a vector that has the following elements:

```{r, echo=FALSE}
desiredVector<-rep(c(2.7,8,3),2)
desiredVector
```
You may ignore the leading comment characters ```##``` and the index notations ```[1]``` appearing at the beginning of the line.

__4.2__Use the ```seq``` function with the ```by``` argument to create a vector that has the following elements:

```{r, echo=FALSE}
desiredVector<-seq(0,2,by=.4)
desiredVector
```

__4.3__ What is returned by the following R commands? (Guess before computing this in R.)

```
primes      <-  c(2,3,5,7,11,13,17,19,23,29)
composites  <-  c(4,6,8,9,10)
primes[composites]
```

__4.4__ What is returned by the following R command?

```
seq(3, 28, by=11) %/% 4
```

__4.5__ Use the ```seq``` function with the ```length.out``` argument to create a vector that has the following elements:

```{r, echo=FALSE}
seq(0, 2, length.out=5)
```

__4.6__ What is returned by the following R commands ?

```
x<-c(2,0,-5,-7)
x[-2.8]
```

__4.7__ What is returned by the following R command ?

```
3 ^ rep(0:2,1:3)
```

__4.8__ Write an R command that uses ```rep``` to create the vector with elements

```{r, echo=FALSE}
rep(1:4,4)+c(rep(0,4),rep(1,4),rep(2,4),rep(3,4))
```

__4.10__ Assign the vector ```x``` to the first eight positive integers. Show that the three R commands

```
x[6:8]
x[c(6:8)]
x[-c(-6:-8)]
```
all yield identical results.

__4.11__ Write an R command that calculates $n^n$ for $n = 1, 2, \cdot \cdot \cdot, 10$.

__Exercises taken from indicated chapter of "Learning Base R", by Lawrence M Leemis, ISBN 978-0-9829174-8-0__

