---
title: "Chapter 8 Answers"
geometry: margin=1.0in
output: pdf_document
---
\pagenumbering{gobble}

__8.1__ Write a one-line function named ```reverse``` with a single arument ```x``` that reverses the order of the elements in the vector ```x```.

```{r}
reverse = function (x) x[length(x):1]
reverse(1:15)
x<-sample(100,10,replace=FALSE)
x
reverse(x)
```

__8.2__ Write a functions namd ```my.cos``` (that calculates the cosine of the supplied angle) with two arguments.
The first argument is named ```angle``` and the second argument is named ```degrees```, which has a default value of ```FALSE```.
Test your function with the R commands 

```
> my.cos(pi/2)
> my.cos(90, degrees = TRUE)
```

```{r}
my.cos = function(angle=NaN, degrees=FALSE) {
    ifelse(is.na(angle),
            NaN,
            ifelse(degrees,
                    cos(angle),
                    cos(180/pi*angle)
            )
    )
}

my.cos(90,TRUE)
my.cos(90,degrees<-TRUE)
my.cos(pi/2)
my.cos()

```

__8.3__  Write a function named ```cube.root``` that calculates the cube root(s) of its argument.
Test your function with the R command
```
> cube.root(c(-8,8,729,1000000))
```

```{r}
-8^(1/3)
# but
cube.root <- function(x) { (x)^(1/3) }
cube.root(c(-8,8,729,1000000))
# and
cube.root <- function(x) { as.numeric(x)^(1/3) }
cube.root(c(-8,8,729,1000000))
# so, ...
cube.root <- function(x) { y<-abs(x)^(1/3) ; ifelse(x>=0,y,-y) ; }
cube.root(c(-8,8,729,1000000))
```

__8.4__ Consider the sample data values $x_1, x_2, \cdot \cdot \cdot, x_n$ and the associated sample order statistics, $x_{(1)}, x_{(2)}, \cdot \cdot \cdot, x_{(n)}$.
The _sample truncated mean_ (also known as the _sample trimmed mean_) is a measure of central tendency defined as 

\begin{center}
$\displaystyle \bar{x} = \frac{x_{(k+1)} + x_{(k+2)} + \cdot \cdot \cdot + x_{(n-k)} }{ n - 2k }$
\end{center}

This is the arithmetic average of the data values with the _k_ lowest and _k_ highest observations removed.
The truncated mean is less sensitive to outliers than the arithmetic mean and is hence known as a _robust estimator_.
This estimator is used in sports that are evaluated by a panel of _n_ judges in which the lowest and highest scores (_k_ = 1) are discarded.
Likewise, the truncated mean is used in calculating the London Interbank Offer Rate (LIBOR) when _n_ = 18 interest rates are collected, and the lowest four and highest four interest rates (_k_ = 4) are discarded.
Assuming that $k < n/2$, write an R function named ```tmean``` with two arguments ```x``` and ```k``` that calculates the sample truncated mean of the elements in the vector ```x``` discarding the ```k``` lowest and ```k``` highest observations.
Test your code with the R commands

```
> tmean(c(9.4, 9.6, 9.1, 9.5, 9.3), 1)
> tmean(1:18, 4)
```


```{r}
tmean = function (x,k) { 
    sorted_x<-sort(x)
    n_from<-k+1
    n_to<-length(x)-k
    mean(sorted_x[n_from:n_to])
}

# case 1
tmean(c(9.4,9.6,9.1,9.5,9.3),1)
# check against :
mean(c(9.4,9.5,9.3))
# case 2
tmean(1:18,4)
# check against :
mean(5:14)
# setup for alternative forms
x<-sample(100,30,replace=FALSE)
x
tmean(x,3)
tmean(x,7)
# alternative form #1
# GOTCHA: parenthesis are REQUIRED on the indexes
tmean = function (x,k) { 
    y<-sort(x)
    mean(y[(k+1):(length(y)-k)])
}

tmean(c(9.4,9.6,9.1,9.5,9.3),1)
tmean(x,3)
tmean(x,7)

# alternative form #2
# sort, then subset, then take the mean
tmean = function (x,k) { mean(sort(x)[(k+1):(length(x)-k)]); }
tmean(c(9.4,9.6,9.1,9.5,9.3),1)
tmean(x,3)
tmean(x,7)
```

__8.5__ Let $x_1, x_2, \cdot \cdot \cdot, x_n$ be the $n > 2$ elements in a vector ```x```.
Write an R function named ```moveave``` with a single argument ```x``` that returns a vector of length $n-1$ whose elements are the moving averages

\begin{center}
$\displaystyle \frac{x_1 + x_2}{2} ,\frac{x_2 + x_3}{2} ,\frac{x_3 + x_4}{2} ,\cdot \cdot \cdot, \frac{x_{n-1} + x_n}{2}$
\end{center}

```{r}
moveave = function (x) { (x[1:length(x)-1] + x[2:length(x)])/2 }
moveave(c(1,2,3))
moveave(c(10,20,30,40,50))
```

__8.6__ Let $x_1, x_2, \cdot \cdot \cdot, x_n$ denote the elements of the vector ```x```.
Write an R function named ```L2``` with a single vector argument ```x``` that calculates the $L_2$ norm

\begin{center}
$\displaystyle \sqrt{x_1^2 + x_2^2 + \cdot \cdot \cdot + x_n^2}$
\end{center}

Test your function with the R commands
```
> L2(c(3,4))
> L2(c(1,1,1))
```

```{r}
L2 = function (x) { sqrt(sum(x^2)) }
L2(c(3,4))
L2(c(1,1,1))
sqrt(3)
```

__8.7__ Let $x_1, x_2, \cdot \cdot \cdot, x_n$ denote the elements of the vector ```x```.
Write an R function named ```Lp``` with a vector argument ```x``` and an integer argument ```p``` that calculates the $p-$norm

\begin{center}
$\displaystyle \Big( \sum_{i=1}^{n} \lvert x_i \rvert ^p \Big)^{(1/p)}$
\end{center}

Test your R function with the R commands

```
Lp(c(3,4),2)
Lp(c(1,1,1),3)
```


```{r}
Lp = function (x, p) { (sum(abs(x)^p))^(1/p)}
Lp(c(3,4),2)
Lp(c(1,1,1),3)
3^(1/3)
Lp(c(1,1,1),2)
```

__8.8__  The built-in R function ```mean``` calculates the sample mean $\bar{x}$. 
Write R functions named ```hmean, gmean,``` and ```qmean``` that calculate the sample harmonic mean, geometric mean, and quadratic mean defined by

\begin{center}
$\displaystyle 
h = \bigg( \frac{1}{n} \sum_{i=1}^{n}\frac{1}{x_i}\bigg)^{-1}, \hspace{0.2in}
g = \bigg( \prod_{i=1}^{n}x_i \bigg)^{1/n}, \hspace{0.2in}
q = \sqrt{\frac{1}{n} \sum_{i=1}^{n} x_i^2 }
$
\end{center}

where $x_1, x_2, \cdot \cdot \cdot, x_n$ are the data values.
Test your functions with a vector of data values of your choice and verify the inequality

\begin{center}
$\displaystyle
min \{ x_1,x_2,\cdot \cdot \cdot,x_n \} \leq h \leq g \leq \bar{x} \leq q \leq max \{ x_1,x_2,\cdot \cdot \cdot,x_n \}
$
\end{center}


```{r}
hmean_orig = function (x) { 1/((1/length(x))*(sum(1/x))) }
hmean_simp = function (x) { length(x) / (sum(1/x)) }
gmean      = function (x) { prod(x)^(1/(length(x))) }
qmean_orig = function (x) { sqrt( (1/length(x)) * (sum(x^2)) ) }
qmean_simp = function (x) { sqrt( (sum(x^2))/length(x) ) }

rv <- sample(100,30,replace = FALSE)
rvmin <-min(rv)
h<-hmean_orig(rv)
g<-gmean(rv)
m<-mean(rv)
q<-qmean_orig(rv)
rvmax<-max(rv)

(rvmin <= h) && (h <= g) && (g<=m) && (m<=q) && (q <= rvmax)
```




### Alternate Ending
```{r}
hmean_alt = function (x,n=length(x)) { n / (sum(1/x)) }
gmean_alt = function (x,n=length(x)) { prod(x)^(1/(n)) }
qmean_alt = function (x,n=length(x)) { sqrt( (sum(x^2))/n ) }

h_alt<-hmean_alt(rv)
g_alt<-gmean_alt(rv)
q_alt<-qmean_alt(rv)

(rvmin <= h_alt) && (h_alt <= g_alt) && (g_alt<=m) && (m<=q_alt) && (q_alt <= rvmax)
```

### Summary
```{r}
# using signif() because digits=4 parameter was ignored
library(knitr)
kable(matrix(
    c(
        c('rvmin',rvmin,rvmin),
        c('h',signif(h),signif(h_alt)),
        c('g',signif(g),signif(g_alt)),
        c('m',signif(m),signif(m)),
        c('q',signif(q),signif(q_alt)),
        c('rvmax',rvmax,rvmax)
    ),
    c(6,3),byrow=TRUE),
    col.names=c('variable','Original','Alternate'),
    align=c('r','l','l'),
    caption="A Comparison of Results")
```

__8.9__ Let $x_1, x_2, \cdot \cdot \cdot, x_n$ be the $n$ elements in the R vector ```x```.
Write an R function named ```mad``` (for _mean absolute deviation_) with a single argument ```x``` that calculates

\begin{center}
$\displaystyle
\frac{1}{n} \sum_{i=1}^{n} \lvert x_i - \bar{x} \rvert
$
\end{center}

where $\bar{x}$ is the sample mean.

_note: this problem includes the $\frac{1}{n}$ correction from author's errata_

```{r}
mad = function (x,x_bar = mean(x),n = length(x) ) { 
    (sum(abs(x-x_bar)))/n 
    }
mad(c(1,2,3))
mad(c(10,20,30,40,50))

```

__8.10__ Write an R function named ```fifth``` with a single vector argument ```x``` that calculates and returns the sample mean of the 5th, 10th, 15th, ... elements of ```x```.
You may assume that ```x``` has at least five elements.

__8.11__ The rectangular coordinate system uses $x$, the signed horizontal distance from the origin, and $y$, the signed vertical distance from the origin, to describe the point $(x, y)$.
The polar coordinate system, on the other hand, uses $r$, the signed distance from the origin, and $\theta$, the signed angle measured counterclockwise from the polar axis, to describe the point $(r, \theta)$.
Assume that $\theta$ is measured in radians.

+   Write an R function named ```polar2rect``` with arguments ```r``` and ```theta``` that returns a two-element vector that contains the rectangular coordinates associated with the point $(r,\theta)$ in the polar coordinate system.
+   Write an R function named ```rect2polar``` with arguments ```x``` and ```y``` that returns a two-element vector that contains the polar coordinates associated with the point $(x,y)$ in the rectangular coordinate system.

__8.12__ Write a one-line function named ```fourth``` that raises its argument to the fourth power.
Then apply the ```formals, body,``` and ```environment``` functions to ```fourth``` in order to isolate the arguments to the function, the R code that comprises the function, and the location of the objects used in the function.


__Exercises taken from indicated chapter of "Learning Base R", by Lawrence M Leemis, ISBN 978-0-9829174-8-0__

