---
title: "Chapter 16 Answers"
geometry: margin=1.0in
output: pdf_document
---
\pagenumbering{gobble}

__Highlights__

+   Review these functions
    +   data.frame
    +   head
    +   tail
    +   summary
    +   transform
    +   subset


__16.1__ What is returned by the follwing R commands? 
(Waking hours from wikipedia.)

```{r}
creatures = c("dog","cat","armadillo","human")
friendly = c(TRUE,TRUE,FALSE,TRUE)
diet = c("cats","mice","termites","Twinkies(tm)")
waking.hours = c(13.9, 11.5, 5.9, 16.0)
creature.data = data.frame (friendly, diet, waking.hours, row.names=creatures)
creature.data


creatures[creature.data$waking.hours < 12 & !creature.data$friendly]
```
__16.2__ Write a single R command that alphabetizes the rows of the data frame ```creature.data``` from Exercise 16.1 by creature name.

```{r}
creature.data[sort(creatures),]
# also (perhaps the separate object creatures was not created)
 creature.data[sort(row.names(creature.data)),]
# for fun:
summary(creature.data)
```

__16.3__ Use the R ```subset``` function to create a data frame consisting of just the creature name and diet associated with friendly creatures who are awake more than 12 hours a day from the data drame ```creature.data``` from Exercise 16.1.

```{r}
subset(creature.data,friendly & (waking.hours > 12),c(diet))
```

__16.4__ Consider the data frame ```creature.data`` from Exercise 16.1.

+   Extract the waking hours for a dog using two different R commands
```{r}
creature.data["dog","waking.hours"]
# subset extracts a new data frame,
# adding the [1,1] indexes extracts the cell value
subset(creature.data,"dog" == row.names(creature.data),c(waking.hours))[1,1]
```

+   Extract the waking hours for all creatures using two different R commands
```{r}
creature.data$waking.hours
creature.data[["waking.hours"]]
subset(creature.data,TRUE,c(waking.hours))[,1]
```

+   Create a data frame that consists only of the rows for dogs and armadillos using two different R commands
```{r}
subset(creature.data,"dog" == row.names(creature.data) | "armadillo" == row.names(creature.data),colnames(creature.data))
subset(creature.data,"dog" == row.names(creature.data) | "armadillo" == row.names(creature.data),)
subset(creature.data,(waking.hours < 16 & waking.hours>12) | waking.hours<6,)
```

+   Create a data frame that consists only of friendly creatures
```{r}
subset(creature.data,friendly,)
```

__16.5__ Consider a data frame named ```a``` with 4 rows and 3 columns populated with positive integers and -1 where -1 denotes a missing value.
Write an R command that replaces each -1 with NA.

```{r}
a = data.frame(col1=1:4,col2=seq(10,40,by=10),col3=seq(100,400,by=100))
a
a[1,1] = a[2,2] = a[3,3] = a$col2[4] = -1
a
a[a==-1] <- NA
a
```

__16.6__ Create the objects ```v```, ```m```, ```a```, ```l```, and ```d``` as a vector, matrix, array, list, and data.frame.
Apply the functions ```class```, ```typeof```, and ```mode ```to each of the objects and summarize the results in a table.

```{r}

v = 1:12
m = matrix(v,3,4)
a = array(v,c(2,2,3))
l = list(v,m,a)
d = subset(creature.data,friendly,)

st = data.frame(
    class  =    c(  class(v),   class(m),   class(a),   class(l),   class(d)), 
    typeof =    c(  typeof(v),  typeof(m),  typeof(a),  typeof(l),  typeof(d)),
    mode   =    c(  mode(v),    mode(m),    mode(a),    mode(l),    mode(d)),
    row.names = c(  "v",        "m",        "a",        "l",        "d")
)
st

```

__16.7__ Create a data frame named ```w``` that consists of the following three named columns:

+   ```x```, the first four positve integers,
+   ```y```, the abbreviations of the first four months,
+   ```z```, the first names of the Beatles.

```{r}
w = data.frame( x = 1:4,
                y = month.abb[1:4],
                z = sort(c("John","Paul","Ringo","George"))
)
w
```

Remove the second column of ```w``` in the following two fashions:

+   set ```w$y``` to ```NULL```, that is, ```w$y = NULL```,
+   subset the columns to keep, that is, ```w[c("x","z")]```.

```{r}
w_ = w
w_$y = NULL

w__ = w
w__[c("x","z")]

```


Show that the first technique alters the data frame ```w ```but the second technique does not alter the data frame ```w```.

```{r}
# the first method was an assignment operation, resulting in a change
w_
# the second method was an extraction operation, with no permanent damage done to the data frame
w__
```


