---
title: "Chapter 15 Answers"
geometry: margin=1.0in
output: pdf_document
---
\pagenumbering{gobble}

__Highlights__

+   Review these functions
    +   list
    +   attributes
    +   unlist

__15.1__ What is returned by the following R commands ?

```{r}
my.list = list(x = 7:2, y = letters )
my.list$y[4]
```

__15.2__ Write a single R command that creates a list named ```x``` with the following three components:

+   the character string "```gnu```",
+   a five-element vector containing the integers 3, 4, 5, 6, 7, and
+   a 12-element vector containing the abbreviations of the names of the months

Write a second R command that extracts the fourth element of the third component.

```{r}
x = list ("gnu" , 3:7 , month.abb)
x[[3]][4]
```

__15.3__ All of the varieties of objects encountered thus far may have names attached to their elements.
Type the following R commands into an R session.

```{r}
x = 1:3
names(x) = c( "first", "second", "third" )
x
y = c(first = 1, second = 2, third = 3)
y
x == y
z = c(first = 1, second = 2, 3)
z
z == x
x == z
z == y
y == z
```

What do you conclude about attaching names to some elements but not to others ?



__Exercises taken from indicated chapter of "Learning Base R", by Lawrence M Leemis, ISBN 978-0-9829174-8-0__

