---
title: "Chapter 10 Answers"
geometry: margin=1.0in
output: pdf_document
---
\pagenumbering{gobble}

__Highlights__

+   Review these functions
    +   complex
    +   Re
    +   Im
    +   Conj
    +   Mod
    +   Arg
+   Don't forget these functions
    +   matrix
    +   rbind
    +   seq
    +   ifelse
    +   ```%%```
    +   ```%/%```

__10.1__ Write a single R command that calculates the product of the complex numbers $2 - 3i$ and $3 + 4i$. 
Check th correctness of the result by hand.

```{r}
(2-3i)*(3+4i)

# or, lets get crazy

my.complex.multiplier = function (c1,c2) {
    # Manual FOIL method
    # Re(c1)*Re(c2) + Re(c1)*Im(c2) + Im(c1)*Re(c2) + Im(c1)*Im(c2)
    # But, Re and Im only return non-imaginary components, so
    # REAL = Re(c1)*Re(c2) * (-1)*(Im(c1)*Im(c2))
    # IMAGINARY = Re(c1)*Im(c2) + Im(c1)*Re(c2)
    return(complex(real = Re(c1)*Re(c2) - Im(c1)*Im(c2),
                    imaginary = Re(c1)*Im(c2) + Im(c1)*Re(c2)))

}

xc1 = complex(real = 2, imaginary = -3)
xc2 = complex(real = 3, imaginary = 4)
my.complex.multiplier(xc1,xc2)
# and confirm:
xc1*xc2

```

__10.2__ What is returned by the following R command? 
Check the correctness of the result by hand.
```
(2 - 3i) ^ 2
```

```{r}
(2-3i)^2
(2-3i)*(2-3i)
my.complex.multiplier(2-3i,complex(real=2,imaginary=-3))
```

__10.3__ Write R commands that calculate $(1-i)^n$, where $i=\sqrt{-1}$, for $n = 4, 8, 16, 20, 24$.
Use the results to write a general mathematical expression for $(1-i)^n$, where $n$ is a multiple of 4.

```{r}
(1-1i)^(seq(4,24,by=4))

n<-seq(4,24,by=4)
ifelse((n%/%4)%%2,-1,1)*4^(n%/%4) + 0i
```

or, 

$\text{Let } m = n \text{ div } 4, then  \\
    \begin{cases} 
      -4^m + 0i & \text{where }m \text{ is odd} \\
       4^m + 0i & \text{where }m \text{ is even} \\
    \end{cases}$


__10.4__ Write R commands to create a 5 x 5 matrix named $w$ whose elements are complex numbers.
The real part of each element is the row number; the imaginary part of each element is the column number.
Compute the matrix that results from multiplying each element of $w$ by its conjugate.

```{r}
seed <- complex(real=1,imaginary=1:5)
w=matrix(rbind(seed,seed+1,seed+2,seed+3,seed+4),5,5)
w
Conj(w)
w*Conj(w)

w[3,3]*Conj(w[3,3])
w[1,2]*Conj(w[1,2])

```



__Exercises taken from indicated chapter of "Learning Base R", by Lawrence M Leemis, ISBN 978-0-9829174-8-0__

