---
title: "Chapter 9 Answers"
geometry: margin=1.0in
output: pdf_document
---
\pagenumbering{gobble}

__Highlights__

+ ```history()```
+ ```history(pattern="some text")```
+ ```options()```
    + digits
    + scipen
    + width
+ ```demo()```
+ ```demo(graphics)```
+ ```??
+ ```appropos()```
+ Notes on Environments
    +   ```search()``` - lists current environments
    +   ```ls(search()[3])``` - lists objects in environment listed by search() at index [3]
    +   ```ls(search()[3],pattern="some text")``` - lists objects in environment listed by search() at index [3] matching "some text"


__9.1__ Use the ```options``` function to set the number of digits displayed to three, the width of the display to 40, and avoid scientific notation.
The use the ```cos``` function to display the cosine of the first ten positive integers to three digits.

```{r}
# setup
targetListOfValues<-c(1:10)
targetListOfValues
# before
cos(targetListOfValues)
# set displayed digits to three
options(digits=3)
cos(targetListOfValues)
# set display width to 40 characters
options(width=40)
cos(targetListOfValues)
# force scientific notation
options(scipen=-1000)
cos(targetListOfValues)
# resist scientific notation
options(scipen=1000)
cos(targetListOfValues)
```

__9.2__ Run the ```help``` function on the ```round``` function.
How many arguments does the ```round``` function accept ?

```{r}
# setup
targetListOfValues<-c(1:10)
options(digits=10)
options(scipen=1000)
options(display=40)
cosValues<-cos(1:10)
cosValues
# round takes two values
#   the vector of values to be rounded
#   signed integer number of decimal places
round(cosValues,2)
round(cosValues,6)
round(cosValues,-2)
round(314.00*cosValues,-2)
```

__9.3__ The three functions (actually _functionals_ because one of their arguments is itself a function) that are useful in numerical problems are 

+   ```uniroot```, which finds the numerical solution to an equation, 
+   ```optimise```, which finds the value associated with the minimum or maximum of a function, and 
+   ```integrate```, which finds the area under a function.

__9.3.a__ Use the ```help``` function to learn about ```uniroot``` and use it to solve $x = e^{-x}$.


```{r}

```

__9.3.b__ Use the ```help``` function to learn about ```optimise``` and use it to gfind the value of the $x$ that minimizes $g(x) = x + e^{-x}$

```{r}

```

__9.3.c__ Use the ```help``` function to learn about ```integrate``` and use it to find the area under the function $h(x) = \frac{sin(x)}{x}$ between $x = 0$ and $x = \pi$, that is

$\int_{0}^{pi} \frac{sin x}{x}dx$

```{r}

```

## Exercise 9.4
```{r}
# confirming from text
system2("bc",stdout=TRUE,stderr="",input="obase=2;12.75")

# as requested in problem
system2("bc",stdout=TRUE,stderr="",input="obase=2;scale=4;19/8")

# confirming as equivalent value
x<-system2("bc",stdout=TRUE,stderr="",input="obase=2;scale=4;19/8")
system2("bc",stdout=TRUE,stderr="",input=paste("obase=10;ibase=2;",x))

#`higlighting the problem identified in the text
system2("bc",stdout=TRUE,stderr="",input="obase=2;scale=4;2/5")

# confirming as approximate value
x<-system2("bc",stdout=TRUE,stderr="",input="obase=2;scale=4;2/5")
system2("bc",stdout=TRUE,stderr="",input=paste("obase=10;ibase=2;",x))

```

## Exercise 9.5
```{r}
system2("bc",stdout=TRUE,stderr="",input="obase=8;ibase=2;101111100")
system2("bc",stdout=TRUE,stderr="",input="obase=16;ibase=2;101111100")
system2("bc",stdout=TRUE,stderr="",input="obase=10;ibase=2;101111100")
```

## Exercise 9.6
```{r}

```

## Exercise 9.7
```{r}

```



__Exercises taken from indicated chapter of "Learning Base R", by Lawrence M Leemis, ISBN 978-0-9829174-8-0__

