---
title: "Chapter 14 Answers"
geometry: margin=1.0in
output: pdf_document
---
\pagenumbering{gobble}

__Highlights__

+   Review these functions
    +   is (testing)
    +   as (coercing)
    

__14.1__ Write an R function named ```Str2Num``` which accepts a single argument which is a character vector whose elements are single digits.
The function should convert this vector of characters to a numeric value.
Test your function with the R command

```
> Str2Num(c("3","6","5"))
```

which should return

```
365
```

```{r}
Str2Num = function (x) {
   # return(sum(10^(rev(1:length(x))-1) * as(x,"numeric")))
    return(
        sum(                    # add up the terms in the vector
            10^(                # compute a "places" scalar
                length(x):1-1   # declining zero-based indexes are powers of 10
            ) 
            *                   # vector multiplication
            as(x,"numeric")     # interpret each string as a number
        )
    )

    # in the test case:
    #   10^2 * 3
    # + 10^1 * 6
    # + 10^0 * 5
}
Str2Num(c("3","6","5"))

```

__14.2__ What is returned by the following R command ?

```{r}
is.character(5 - 3i)
```

__14.3__ What is returned by the following R command ?

```{r}
x = 3 + 4i
as.integer(x * Conj(x))
```

__14.4__ Write an R function named ```benford``` (after Benford's Law) that returns the leading digit of its single numeric argument.
The leading digit of 365 is 3; the leading digit of 0.0243 is 2.

```{r, quiet=TRUE, warnings=FALSE}
benford = function(x) {
    y <- abs(x)                                 # eliminate negative signs
    while ( y < 1 ) y <- y * 10                 # eliminate leading zeroes
    return(
        suppressWarnings(                       # eliminate R messages about decimals
            as.numeric(                         # convert back to numeric
                unlist(                         # reverse effect of strsplit
                    strsplit(                   # make each digit a separate character in a vector
                        as.character(y),''      # turn the number into a character string
                    )
                )
            )[1]                                # take the first vector element
        )
    )
}
benford(365)
benford(0.0243)
benford(-17)
benford(-pi/1000)


benford_ = function(x) {
    y <- abs(x)                         # eliminate negative numbers
    while ( y < 10  ) y <- y * 10       # eliminate leading zeroes
    while ( y > 100 ) y <- y / 10       # setup for integer division
    return ( y %/% 10)                  # integer division for result
}

benford_(365)
benford_(0.0243)
benford_(-17)
benford_(-pi/1000)

benford__ = function(x) {
    y  <- unlist(strsplit(as.character(x),'')) 
    z  <- ifelse(y=="-","0",ifelse(y==".","0",y))
    zz <- as.numeric(z)
    return(zz[zz>0][1])
}

benford__(365)
benford__(0.0243)
benford__(-17)
benford__(-pi/1000)

benford___ = function(x) {
    x <- abs(x)                 # deal with negative numbers
    while(x < 1) x <- x * 10    # deal with smallish numbers
    return(
        as.numeric(             #       convert back to numeric
            substr(             #   **new** pick off first character
                as.character(x) # convert number to character vector, length=1
                ,1              # start = 1
                ,1              # stop = 1
            )
        )
    )
}

benford___(365)
benford___(0.0243)
benford___(-17)
benford___(-pi/1000)
```

__14.5__ Guess the effect of automatic coercion in the following R commands:

```{r}
c(0,TRUE)
c(2,FALSE)
c(2L, 5)
c(3L, 4.4)
c(2.3, NA)
c(3, "R")
c(FALSE, 9L, 4.4, "S")

class(c(0,TRUE))
class(c(2,FALSE))
class(c(2L, 5))
class(c(3L, 4.4))
class(c(2.3, NA))
class(c(3, "R"))
class(c(FALSE, 9L, 4.4, "S"))
```
