---
title: "Chapter 13 Answers"
geometry: margin=1.0in
output: pdf_document
---
\pagenumbering{gobble}

__Highlights__

+   Review these functions
    +   ifelse 
    +   sqrt
    +   round
    +   seq
    +   matrix

__13.1__ Let ```x``` be a vector containing numerical elements, each of which is a positive integer.
Write a single line of R code that replaces each element that is a perfect square with zero.

```{r}
x <- 1:100
x <- ifelse((round(sqrt(x)) == sqrt(x)),0,x)
x
```

__13.2__ What is returned by the following R commands?

```{r}
month.len = c(31,28,31,30,31,30,31,31,30,31,30,31)
month.abb[month.len < 30]
```

__13.3__ What is returned by the following R commands?

```
> x = seq(1,9, length.out=5)
> y = -1:3
> x[x <= y ^ 2]
```

```{r}
x = seq(1,9, length.out=5)
x
y = -1:3
y
x <= y ^ 2
x[x <= y ^ 2]
```

__13.4__ Suppose that ```x``` is a vector of numeric elements.
Write R commands that create a vector ```y``` of character elements that assume the character strings "positive", "zero", or "negative" associated with the corresponding values of ```x```. 
For example, the ```y``` vector associated with ```x = c(3, 5, 0, -7) ``` is ```y = c("positive","positive","zero","negative")```.

```{r}
x <- c(3, 5, 0, -7) 
y <- ifelse(x==0,"zero",ifelse(x<0,"negative","positive"))
y
```

__13.5__ Let ```x``` be a vector that consists of numeric elements.
Write a single R command to find the mean of the positive elements of ```x```.

```{r}
x <- c(-100,-100,-100,10,20,30,40,50)
mean(x)
mean(x[x>0])
x[x>0]
```

__13.6__ Let ```x``` be a vector that consists of numeric elements.
Write a single R command that returns a vector containing the subscripts of the positive elements of ```x```.

```{r}
x <- c(-100,-100,-100,10,20,30,40,50)
(1:length(x))[x>0]
```


__13.7__ Write an R function named ```DoubleTriple``` with a single argument ```mx```, which is a matrix that consists of integer-value elements.
This function should return a matrix with the same dimensions as ```mx```, but with odd-valued elements in ```mx``` doubled and even-valued elements in ```mx``` tripled.

```{r}
DoubleTriple = function(mx) {
    return ( ifelse( (mx %% 2) == 1, mx*2, mx*3) )
}   

mx_input <- matrix(1:2,12,12)
mx_input
y <- DoubleTriple(mx_input)
y
```


__13.8__ What is returned by the following R commands ?

```{r}
A = matrix(1:15, nrow=3)
mean(A[1:2,3:4] > 10)
```



__Exercises taken from indicated chapter of "Learning Base R", by Lawrence M Leemis, ISBN 978-0-9829174-8-0__

