# Exercises, Chapter 6

## Exercise 6.1

True or False: A matrix is a special case of an array.

+ text, p.32, 'The ```array``` function is a generalization of the ```matrix``` function.'
+ A _matrix_ is an _array_ whose third and final subscript has a value of one.
+ In my notes, I have the phrase "single layer" written next to this question.

## Exercise 6.2

Write an R command that assigns the first 18 positive integers to the 3 x 3 x 2 array named _x_ using the default input convention.
Next, write an R command that extracts the elements of _x_ in a fashion that includes the second and third rows, excludes the second column, and includes the second layer.

```{r}
# x<-array()        create array, x, 
# seq(1:18)         populated with first 18 positive integers
# dim<-c(3,3,2)     with dimensions 3x3x2
#                   using default input convention
x<-array(seq(1:18),dim<-c(3,3,2))
x

# x[]   extract elements of x
# 2:3   include second and third rows
# -2    exclude second column
# 2     include second layer
x[2:3,-2,2]

```

## Exercise 6.3

The 3 x 4 x 5 array _a_ is filled with numeric values.
The R command ```> b = a[c(1,3),-c(1,4),3:5]``` sets _b_ to a 2 x 2 x 3 array that includes the first and third rows, excludes the first and fourth columns, and includes the third, fourth, and fifth layers of _a_. 
Write an R command that accomplishes this same operation using the smallest number of keystrokes.


```{r}
a<-array(seq(1:60),dim<-c(3,4,5))
a[c(1,3),-c(1,4),3:5]
a[-2,2:3,3:5]
```

