# Exercises, Chapter 5

## Exercise 5.1

Using the _minimum number of keystrokes,_ write a single R command that constructs the matrix _A_ given below.

_A_ = 
```{r, echo=FALSE}
rbind(5:9,rep(12,5),(9:5)^2,c(2,8,4,7,3))
```

## Exercise 5.2

Write R commands that create a 2 x 4 matrix named _x_ that contains the inteegers 11, 12, ..., 18 arranged in a column-wise fashion.
Then use subscripts that extact the elements that exclude the first row and include the first and fourth columns.

```{r, echo=FALSE}
x<-matrix(11:18,2,4)
x
```
to get 
```{r, echo=FALSE}
x[-1,c(1,4)]
```

## Exercise 5.3

Let A and B be the following 3 x 2 matrices:


```{r, echo=FALSE}
A<-matrix(1:6,3,2)
B<-A+6
```
A=
```{r, echo=FALSE}
A 
```
B=
```{r, echo=FALSE}
B 
```

```{r, echo=FALSE}
# rbind(A,B)
# cbind(A,B)
```

Write R commands to create the matrices _A_ and _B_ using the ```matrix``` function, then combine _A_ and _B_ to create two new matrices with the ```rbind``` and ```cbind``` functions.

## Exercise 5.4

A _one-step transition matrix_ has elements that are between 0 and 1, row sums that equal 1, and the same number of rows and columns.
Write an R command to create the 2 x 2 one-step transition matrix _A_ with elements given by

```{r, echo=FALSE}
# easier to read and infer
# A<-c(0.8,0.3)
# cbind(A,1-A)
# or, single command; may expect values in objects
cbind(c(0.8,0.3),1-c(0.8,0.3))
```

## Exercise 5.5

Write a single R command that creates a sequence of one 1, two 2s, three 3s, and four 4s, and places them in a 2 x 5 matrix named ```b``` in a row-wise fashion.
This command should use as few keystrokes as possible.

```{r, echo=FALSE}
# matrix(rep(1:4,1:4),2,5,T)
```

